---
title: "ML assignment unit 3"
author: "saravanan s"
date: "17/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

---

```{r}
library(ISLR)
library(MASS)
attach(Default)
```

1. (a) Create a binary variable, mpg01, that contains a 1 if mpg contains a value above its median, and a 0 if mpg contains a value below its median. You can compute the median using the median() function. Note you may find it helpful to use the data.frame() function to create a single data set containing both mpg01 and the other Auto variables.

```{r}
attach(Auto)
mpg01 <- rep(0, length(mpg))
mpg01[mpg > median(mpg)] <- 1 
Auto_mpg01 <- data.frame(Auto, mpg01)
dim(Auto_mpg01)
head(Auto_mpg01[, -9])
attach(Auto_mpg01)
```

  Ans : 
      Create a binary variable name is called mgp01 , use  median() function, and combine the binary variable mpg01 and Auto as a Single data set 

(b) Explore the data graphically in order to investigate the association between mpg01 and the other features. Which of the other features seem most likely to be useful in predicting mpg01? Scatterplots and boxplots may be useful tools to answer this question. Describe your findings

```{r}
cor(Auto_mpg01[, -9]) # Show the correlations between the variables
cols <- character(nrow(Auto_mpg01))
cols[] <- "black"
cols[Auto_mpg01$mpg01 == 1] <- "orange"
cols[Auto_mpg01$mpg01 == 0] <- "black"
pairs(Auto_mpg01, col=cols) # Plot the scatterplot matrix
```
```{r}
par(mfrow=c(2,3))
boxplot(weight ~ mpg01, data = Auto_mpg01, main = "Weight", 
        xlab = "mpg01", ylab = "Weight",
        col = c("blue", "red"))
boxplot(year ~ mpg01, data = Auto_mpg01, main = "Year", 
        xlab = "mpg01", ylab = "Year",
        col = c("red", "blue"))
boxplot(cylinders ~ mpg01, data = Auto_mpg01, main = "Cylinders", 
        xlab = "mpg01", ylab = "Cylinders",
        col = c("blue", "red"))
boxplot(acceleration ~ mpg01, data = Auto_mpg01, main = "Acceleration", 
        xlab = "mpg01", ylab = "Acceleration",
        col = c("blue", "red"))
boxplot(displacement ~ mpg01, data = Auto_mpg01, main = "Displacement", 
        xlab = "mpg01", ylab = "Displacement",
        col = c("blue", "red"))
boxplot(horsepower ~ mpg01, data = Auto_mpg01, main = "Horsepower", 
        xlab = "mpg01", ylab = "Horsepower",
        col = c("blue", "red"))
```

  Ans : As we have observed from the above plot that  `mpg01` has not correlation between these variables `Weight`, `Cylinders`, `Displacement`, and `Horsepower`.

(c) Split the data into a training set and a test set.
```{r}
train <- (year %% 2 == 0)
Auto_mpg01.train <- Auto_mpg01[train, ]
Auto_mpg01.test <- Auto_mpg01[!train, ]
mpg01.test <- mpg01[!train]
```
  Ans :  splitted the dataset  into   training set and a test set
       training data have 210 records
       test data have 182 records
        
  
(d)	Perform LDA on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained? 

```{r}
fit.lda <- lda(mpg01 ~ cylinders + weight + displacement + horsepower,data = Auto_mpg01, subset = train)
fit.lda
plot(fit.lda)
lda.class <- predict(fit.lda, Auto_mpg01.test)$class
table(lda.class, mpg01.test)
mean(lda.class != mpg01.test)
```
  Ans : Using all 4 predictors (`cylinders`, `weight`, `displacement`, and `horsepower`) with a Linear Discriminant Analysis the test error rate is 12.63736%.


(e)	Perform QDA on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained? 

```{r}
qda.fit <- qda(mpg01 ~ cylinders + weight + displacement + horsepower, 
               data = Auto_mpg01, subset = train)
qda.fit
qda.class <- predict(qda.fit, Auto_mpg01.test)$class
table(qda.class, mpg01.test)
mean(qda.class != mpg01.test)
```
  Ans : Using the all 4 predictors (`cylinders`, `weight`, `displacement`, and `horsepower`) with a Quadratic Discriminant Analysis the test error rate is 13.18681%

(f)	Perform logistic regression on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?

```{r}
glm.fit <- glm(mpg01 ~ cylinders + weight + displacement + horsepower, data = Auto_mpg01, family = binomial, subset = train)
summary(glm.fit)
probs <- predict(glm.fit, Auto_mpg01.test, type = "response")
glm.pred <- rep(0, length(probs))
glm.pred[probs > 0.5] <- 1
table(glm.pred, mpg01.test)
mean(glm.pred != mpg01.test)
```
  Ans : Using all 4 predictors (`cylinders`, `weight`, `displacement`, and `horsepower`) with a Logistic Regression Analysis the test error rate is 12.08791%.

(g)	Perform KNN on the training data, with several values of K, in order to predict mpg01. Use only the variables that seemed most associated with mpg01 in (b). What test errors do you obtain? Which value of K seems to perform the best on this data set?

```{r}
library(class)
train.X <- cbind(cylinders, weight, displacement, horsepower)[train, ]
test.X <- cbind(cylinders, weight, displacement, horsepower)[!train, ]
train.mpg01 <- mpg01[train]
###  K = 1
set.seed(12345)
knn.pred.1 <- knn(train.X, test.X, train.mpg01, k = 1)
table(knn.pred.1, mpg01.test)
mean(knn.pred.1 != mpg01.test)
```

  Ans : With KNN Analysis (K = 1), the test error rate is `r mean(knn.pred.1 != mpg01.test) * 100`%

```{r}
###  K = 5
set.seed(12345)
knn.pred.5 <- knn(train.X, test.X, train.mpg01, k = 5)
table(knn.pred.5, mpg01.test)
mean(knn.pred.5 != mpg01.test)
```
With KNN Analysis (K = 5), the test error rate is `r mean(knn.pred.5 != mpg01.test) * 100`%
